---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# bis620

<!-- badges: start -->
[![R-CMD-check](https://github.com/hugoclho/bis620/workflows/R-CMD-check/badge.svg)](https://github.com/hugoclho/bis620/actions)
<!-- badges: end -->

The goal of `bis620` is to build a package for Yale course BIS 620.

## Installation

You can install the released version of bis620 from [CRAN](https://CRAN.R-project.org) with:

``` r
install.packages("bis620")
```

And the development version from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("hugoclho/bis620")
```
## Example

This is a basic example which shows you the overview of the package `bis620`

```{r example}
library(bis620)
## basic example code
```

This package includes data sets `ae`, `aeendpt`, `asendpt` from study `NCT00460265`

```{r cars}
head(ae[,1:8])
```
Using build in fuction `get_na()` we can examine the number of NA of each column in the data set:

```{r pressure, echo = FALSE}
get_na(ae)
```

Using build in function `get_min_na()` / `get_max_na()` we can examine the column with min/max number of NA in the data set:

```{r}
get_min_na(ae)
get_max_na(ae)
```


## Sparse Matrix Class `bis620_sparse_matrix`

Create bis620_sparse_matrix object

```{r}
x1 <- bis620_sparse_matrix(
  i = c(1, 2, 5, 6),
  j = c(2, 2, 6, 1),
  x = c(4.3, 5.6, 7, 10)
)
```

Print bis620_sparse_matrix object like other normal R matrices

```{r}
print(x1)
```

Operations between bis620_sparse_matrix returns a bis620_sparse_matrix, and `+`, `-`, `/`, `*` along with `%*%` are compatible with other types of R matrices

```{r}
x1 + x1
print(x1 + x1)
```


```{r}
# compatibility with other Matrix types
library(Matrix)
x2 <- matrix(rnorm(36), ncol = 6)

set.seed(1)

x3 <- Matrix::Matrix(rnorm(36), ncol = 6)

x4 <- Matrix::sparseMatrix(
  i = c(1, 1, 3, 6),
  j = c(2, 3, 5, 1),
  x = c(4.3, 5.6, 7, 10),
  dims = c(6, 6)
)
```


```{r}
x1 + x1
print(x1 + x1)
x1 - x4
t(x1) %*% x1
print(t(x1) %*% x1)
x1 %*% x1
x1 + t(x2)
x1 %*% x2
x3 + x1
x3 %*% x3
x1 + x4
x1 %*% x4
```

## Final Homework Proposal
Find a real world data set. Call Python in R and finish data cleaning and exploration under Python. Import data to R and create R6 object and related methods. 










