---
title: "homework-3"
authur: Hugo He, Arthur Wang
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{homework-3}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup,, message=FALSE}
library(bis620)
library(Matrix)
```


However, there is a lot of room for improvement with this implementation. In this homework, you may work in groups of up to 3. Please do the following:

1. Create a sparse matrix class that either holds data with the `data.frame` representation above or is a `data.frame` with the above representation. (1 points)
2. Add arithmetic operators `+`, `-`, `/`, and `*` along with `%*%` so that the matrix behaves like a regular R matrix. (2 points)
3. Add a print function so that it's easier to see the layout. You may use the `Matrix::sparseMatrix` print method if it's helpful. (1 point)
4. Make sure that your matrix is compatible with other R matrics. Use the code below to test it. (2 points)
5. Add the new matrix, the operators, and the tests into your `bis620` package. (2 points)
6. Propose a final homework for this class. (2 points)



```{r eval = TRUE}
# Create matrices for testing
x1 <- bis620_sparse_matrix(
  i = c(1, 2, 5, 6), 
  j = c(2, 2, 6, 1), 
  x = c(4.3, 5.6, 7, 10)
)

set.seed(1)

x2 <- matrix(rnorm(36), ncol = 6)

set.seed(1)

x3 <- Matrix(rnorm(36), ncol = 6)

x4 <- sparseMatrix(  
  i = c(1, 1, 3, 6), 
  j = c(2, 3, 5, 1), 
  x = c(4.3, 5.6, 7, 10),
  dims = c(6, 6)
)
```

## 1. `+` operations

```{r}
# 1. + operations
# Test that the following are correct:
x1 + x1
print(x1 + x1)
x1 + t(x2)
x3 + x1
x1 + x4
```

## 2. `-` operations

```{r}
# 2. - operations
# Test that the following are correct:
x1 - x1
x1 - t(x2)
x3 - x1
x1 - x4
```

## 3. `*` operations

```{r}
# 3. * operations
# Test that the following are correct:
x1 * x1
print(x1 * x1)
x1 * t(x2)
x3 * x1
x1 * x4
```

## 4. `/` operations

```{r}
# 4. / operations
# Test that the following are correct:
x1 / x1
print(x1 / x1)
x1 / t(x2)
x3 / x1
x1 / x4
```

## 5. `%*%` operations

```{r}
# 5. %*% operations
# Test that the following are correct:
t(x1) %*% x1
print(t(x1) %*% x1)
x1 %*% x1
print(x1 %*% x1)
x1 %*% x2
x3 %*% x3
x1 %*% x4
```

## 6. Print sparse matrix: 

```{r}
x1
print(x1)
```


## 7. All original test cases: 

```{r}
# all original test cases: 
x1 + x1
print(x1 + x1)
x1 - x1
t(x1) %*% x1
print(t(x1) %*% x1)
x1 %*% x1
x1 + t(x2)
x1 %*% x2
x3 + x1
x3 %*% x3
x1 + x4
x1 %*% x4
```

